---
title: "EALS Study 2 analysis"
author: "Mike Frank"
date: "2024-02-26"
output: html_document
---

```{r}
library(tidyverse)
library(here)
library(lme4)
```

Read data. 

```{r}
d <- read_csv(here("study2","EALS_study2_concatenated_allwaves.csv"))
```

Descriptives.

```{r}
d |> 
  group_by(datawave, condition) |>
  summarise(mean_age = mean(age_m/12), 
            prop_male = mean(gender_male1 == "Male"), 
            n = n()) |>
  knitr::kable(digits = 2)
```
# Plot condition effects

Let's not do cool survival curves and just look at the simple first bid time.

```{r}
ggplot(d, aes(x = condition, y = first_bid_toy_s)) + 
  geom_jitter(height = 0, width = .2) + 
  stat_summary(fun.data = "mean_cl_boot", col = "red") + 
  facet_wrap(~datawave) +
  scale_y_log10() + 
  ylab("First bid (s)")
```


```{r}
ggplot(d, aes(x = condition, y = bid_help_toy)) + 
  geom_jitter(height = .05, width = .2) + 
  stat_summary(fun.data = "mean_cl_boot", col = "red") + 
  facet_wrap(~datawave) +
  ylab("Bids for help")
```


# Exploration

Relations between exploration time and demographics. 

```{r}
ggplot(d, aes(x = age_m, y = first_bid_toy_s, col = datawave)) + 
  geom_point() + 
  geom_smooth(method = "lm", aes(group = 1), col = "black", lty = 2) + 
  geom_smooth(method = "lm", se = FALSE) + 
  scale_y_log10() + 
  ylab("First bid (s)")
```

Add gender. 

```{r}
ggplot(d, aes(x = age_m, y = first_bid_toy_s, col = gender_male1)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = TRUE) + 
  scale_y_log10() + 
  ylab("First bid (s)")
```

# Models

Trying some kitchen sink models to understand data. 

```{r}
summary(lm(log(first_bid_toy_s) ~ condition, 
             data = d))
```

Add age. 

```{r}
summary(lm(log(first_bid_toy_s) ~ condition * age_m, 
             data = d))
```
Try a mixed effects model. This model doesn't converge because not all waves have all conditions, but it does show some condition effect. 

```{r}
summary(lmer(log(first_bid_toy_s) ~ condition + (condition | datawave), 
             data = d))
```

